\section*{About the Instructor}
\textbf{Name:} \bluelink{mailto:sarthok@iiitd.ac.in}{Sarthok Sircar}

\textbf{Office Hours:} By appointment

\section*{About the Course}
\textbf{Course Code and Name:} MTH203: Multivariate Calculus (Math-III)

\textbf{Prerequisites:} None

\textbf{Anti-requisites:} MTH240: Real Analysis-I

\textbf{Description:}

This course covers topics in multivariable calculus, vector calculus, and complex analysis. The course starts with an
extension of concepts like limits, continuity, and differentiation to functions of multiple variables. Partial
differentiation is defined and applied. The idea of Taylor series is extended to functions of two variables. This is
followed by a definition and application of integration in more than one dimensions, i.e. double and triple integrals. We
then work with vector functions and develop the ideas of vector fields and differentiation and integration as applicable to
vector calculus (gradient, divergence, curl, line and surface integrals, etc.) The ideas of circulation and flux are
developed and Green's, Stoke's and divergence theorems are covered. The final module of the course deals with complex
numbers and functions and an extension of concepts of calculus to complex variables.

\textbf{Grading Scheme:}
\begin{enumerate}
    \item \textbf{30\%}: 12 Assignments/Worksheets to be solved in tutorials. Out of 12, best 10 assignments will be
    considered for grading, making each assignment worth 3\%.

    \item \textbf{35\%}: 1 Mid-Sem Exam.
    \item \textbf{35\%}: 1 End-Sem Exam.
\end{enumerate}

\textbf{Important Links:}
\begin{enumerate}
    \item \bluelink{http://techtree.iiitd.edu.in/viewDescription/filename?=MTH203}{Course Directory}
    \item \bluelink{https://classroom.google.com/c/NDA4NDczODQwNTla}{Google Classroom}
\end{enumerate}

\pagebreak